prolfqua - An R-package for Analysing Proteomics Label Free Quantification Experiments

### 2022-03-18 inst/MyArticle/paper.Rmd
Edits on code snippets

Peptide intensities are $\log_2$ transformed and robust z-score scaled but then rescaled to preserves the original range of the data using the method `robscale`. We show the distribution of the normalized intensities in Figure \@ref(fig:prepro) Panel A. Protein intensities are estimated using Tukey's median polish. Figure \@ref(fig:prepro) Panel B shows the peptide intensities and the estimated protein intensities. The boxplot (Panel D) shows the distribution of the protein intensity estimate in each condition. Finally, Panel C shows the distribution of the protein standard deviations for each condition.

```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.width = 8, fig.height = 8}
library(prolfqua)
d <- prolfqua_data('data_ionstar')$filtered()
lfqd <- LFQData$new(d$data, d$config) 
t <- lfqd$get_Transformer()
lfqd <-  t$log2()$robscale()$lfq
pl <- lfqd$get_Plotter()
p_a <- pl$intensity_distribution_density() +
  ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
agr <- lfqd$get_Aggregator()
lfqp <- agr$medpolish()

p_b <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
p_c <- lfqp$get_Stats()$violin() + ggplot2::labs(tag = "C")
pl <- lfqp$get_Plotter()
p_d <- pl$boxplots()$boxplot[[1]] + ggplot2::labs(tag = "D")
ggpubr::ggarrange(p_a, p_b, p_c, p_d)

@@

```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%'}
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
)
# fit model
lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
modelFunction <- strategy_lm( lmmodel, model_name = "lm")
models <- build_model(lfqp, modelFunction)
p0 <- models$anova_histogram()
# compute contrasts from linear model and with imputation
contr <- prolfqua::Contrasts$new(models, contrasts) |> 
  prolfqua::ContrastsModerated$new()
conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) |> 
  prolfqua::ContrastsModerated$new()
tmp <- prolfqua::addContrastResults(prefer = contr, add = conI)
# visualize results
pl <- tmp$merged$get_Plotter()
p1 <- pl$histogram()$p.value
p2 <- pl$volcano()$FDR +
   ggplot2::theme(legend.position = "bottom")
p1
gridExtra::grid.arrange(p1, p2, ncol = 2)

@@

The peptide intensities from the _peptide.txt_ file were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only. We removed _one hit wonders_, i.e. proteins with a single peptide assignment. No further data filtering was used. We also did not use imputation. The protein level linear model was fitted to the protein intensities inferred from the peptide intensities using the Tukey's median polish. The mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model were fitted to the peptide intensities. The details of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` benchmark implementation are available online  [BenchmarkingIonstarData](https://wolski.github.io/prolfqua/articles/BenchmarkingIonstarData.html).

@@

# Supplement

>>>>

The code snippets in this section demonstrate the basic usage of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package.
Peptide intensities are $\log_2$ transformed and robust z-score scaled but then rescaled to preserves the original range of the data using the method `robscale`. We show the distribution of the normalized intensities in Figure \@ref(fig:prepro) Panel A. Protein intensities are estimated using Tukey's median polish. Figure \@ref(fig:prepro) Panel B shows the peptide intensities and the estimated protein intensities. The boxplot (Panel D) shows the distribution of the protein intensity estimate in each condition. Finally, Panel C shows the distribution of the protein standard deviations for each condition.

```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.width = 8, fig.height = 8}
## load peptide level data
d <- prolfqua::prolfqua_data('data_ionstar')$filtered()

## create R6 obejct
lfqd <- prolfqua::LFQData$new(d$data, d$config)

##  transform intensities
t <- lfqd$get_Transformer()

lfqd <-  t$log2()$robscale()$lfq

## plotting
pl <- lfqd$get_Plotter()
p_a <- pl$intensity_distribution_density() +
  ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")

## infer protein intensities from peptide intens
agr <- lfqd$get_Aggregator()
lfqp <- agr$medpolish()

## plotting
p_b <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
p_c <- lfqp$get_Stats()$violin() + ggplot2::labs(tag = "C")
pl <- lfqp$get_Plotter()
p_d <- pl$boxplots()$boxplot[[1]] + ggplot2::labs(tag = "D")
ggpubr::ggarrange(p_a, p_b, p_c, p_d)
```

@@

```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%'}
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
)
## fit model
lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
modelFunction <- prolfqua::strategy_lm( lmmodel, model_name = "lm")
models <- prolfqua::build_model(lfqp, modelFunction)
p0 <- models$anova_histogram()

## compute contrasts from linear model and with imputation
contr <- prolfqua::Contrasts$new(models, contrasts) |> 
  prolfqua::ContrastsModerated$new()

conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) |> 
  prolfqua::ContrastsModerated$new()

# TODO(wew): give tmp a meaningful name
tmp <- prolfqua::addContrastResults(prefer = contr, add = conI)

## visualize results
pl <- tmp$merged$get_Plotter()
p1 <- pl$histogram()$p.value
p2 <- pl$volcano()$FDR +
   ggplot2::theme(legend.position = "bottom")

gridExtra::grid.arrange(p1, p2, ncol = 2)
```

@@

The peptide intensities from the MaxQuant _peptide.txt_ file were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only. We removed _one hit wonders_, i.e. proteins with a single peptide assignment. No further data filtering was used. We also did not use imputation. The protein level linear model was fitted to the protein intensities inferred from the peptide intensities using the Tukey's median polish. The mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model were fitted to the peptide intensities. The details of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` benchmark implementation are available online  [BenchmarkingIonstarData](https://wolski.github.io/prolfqua/articles/BenchmarkingIonstarData.html).

@@

# Supplements

## Creating a prolfqa configuration

The following code snipet demomnstarte the usage from MSFragger ouput.
